---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, echo=TRUE}

#read data
active_data <- read.csv("activity.csv")

#convert variable "date"'s attribute from "factor" to "Date"
active_data$date <- as.Date(as.character(active_data$date))


```

## What is mean total number of steps taken per day?

```{r, echo=TRUE}

library(ggplot2)

# split data by date
data_by_date <- split(active_data, active_data$date)

steps_per_day <- sapply(data_by_date, function(x) sum(x$steps, na.rm = TRUE))

# make data tidy
steps_per_day <- data.frame(date = names(steps_per_day), steps=steps_per_day, row.names = NULL)

mean_steps_per_day <- mean(steps_per_day$steps)

median_steps_per_day <- median(steps_per_day$steps)

qplot(steps_per_day$steps, geom="histogram") + xlab("steps")


```

Mean of totl number of steps per day is  `r mean_steps_per_day`

Median of totl number of steps per day is  `r median_steps_per_day`

## What is the average daily activity pattern?

```{r, echo=TRUE}

#turn off scientific notation
options(scipen=999)

five_interval <- split(active_data, active_data$interval)

steps_per_interval <- sapply(five_interval, function(x) mean(x$steps, na.rm = TRUE))

# make data tidy
steps_per_interval <- data.frame(interval = names(steps_per_interval), steps=steps_per_interval, row.names = NULL)
 
steps_per_interval$interval <- as.numeric(as.character(steps_per_interval$interval))

# draw  time series plot
plot(steps_per_interval)

index_of_max_steps <- which(steps_per_interval$steps == max(steps_per_interval$steps))


```

For `r  steps_per_interval[index_of_max_steps, ]$interval` interval have max mean steps of  `r steps_per_interval[index_of_max_steps, ]$steps`


## Imputing missing values

There are `r sum(is.na(active_data))` NA value in dataset.

```{r, echo=TRUE}

# define a function
guess_steps_by_interval <- function (interval_value)  {
  
  return (steps_per_interval[which(steps_per_interval$interval == interval_value), ]$steps)
}
 
# copy data to a new data to be process
new_data <- active_data
 
# fill NA with new value
for(i in 1:nrow(new_data)) {
  
  if(is.na(new_data[i, ]$steps) == TRUE) {
    
    new_data[i, ]$steps <- round(guess_steps_by_interval(new_data[i, ]$interval))
  }
  
}


# split data by date
data_by_date_new <- split(new_data, new_data$date)

steps_per_day_new <- sapply(data_by_date_new, function(x) sum(x$steps, na.rm = TRUE))

# make data tidy
steps_per_day_new <- data.frame(date = names(steps_per_day_new), steps=steps_per_day_new, row.names = NULL)

mean_steps_per_day_new <- mean(steps_per_day_new$steps)

median_steps_per_day_new <- median(steps_per_day_new$steps)

qplot(steps_per_day_new$steps, geom="histogram") +  xlab("steps")

```

Mean of totl number (NA processed) of steps per day is  `r mean_steps_per_day_new` while old mean is `r mean_steps_per_day`

Median of totl number (NA processed) of steps per day is  `r median_steps_per_day_new` while old median is  `r median_steps_per_day`

## Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}
library(lubridate)

#Change Language to English
Sys.setlocale("LC_TIME", "English")
 
# Add a new column indicate is this day a weekend or
new_data$day <- ifelse((weekdays(new_data$date) == "Sunday" ), 
                       "weekend", "weekday")

new_data$day <- ifelse((weekdays(new_data$date) == "Saturday"), 
                       "weekend", "weekday")

new_data$day <- as.factor(new_data$day)

# prepare compare plot

data_by_weekday <- split(new_data, new_data$day)




caculate_interval_patern <- function(d) {
  
  fi <- split(d, d$interval)

  steps_per_interval_local <- sapply(fi, function(x) mean(x$steps, na.rm = TRUE))
  
  # make data tidy
  steps_per_interval_local <- data.frame(interval = names(steps_per_interval_local), 
                                         steps=steps_per_interval_local, row.names = NULL)
   
  steps_per_interval_local$interval <- as.numeric(as.character(steps_per_interval_local$interval))
  
  return(steps_per_interval_local)
  
}

weekday_patern <- caculate_interval_patern(data_by_weekday$weekday)

weekend_patern <- caculate_interval_patern(data_by_weekday$weekend)

# plot differnt
par(mfrow = c(2,1))
plot(weekday_patern, type = "l", xlab = "interval", ylab= "Number of steps", main = "weekday")
plot(weekend_patern, type = "l", xlab = "interval", ylab= "Number of steps", main = "weekend")

```






